%----------------------------------------------------------
% Number of pages for Report 25-40
%----------------------------------------------------------
% List of commands to convert synopsis.tex to synopsis.pdf
%----------------------------------------------------------
%pdflatex synopsis.tex
%evince synopsis.pdf &
\documentclass[12pt,singleside,a4paper]{article}
\usepackage{epsfig}
\usepackage{cite}
\usepackage{geometry}
\geometry{left=20mm,right=20mm,top=20mm,bottom=50mm}
\pagestyle{plain}
\begin{document}
\pagenumbering{arabic}
\input{be_title.tex}
\input{be_certificate2.tex}
\section*{\centering{Declaration}}

\vspace{3.5cm}
I wish to state that the work embodied in this work titled “project title” forms my own contribution to the work carried out under the guidance of Guide Name at the Sardar Patel Institute of Technology. I declare that this written submission represents my ideas in my own words and where others' ideas or words have been included, I have adequately cited and referenced the original sources. I also declare that I have adhered to all principles of academic honesty and integrity and have not misrepresented or fabricated or falsified any idea/data/fact/source in my submission.
\\
\\
\\
\textbf{Candidate 1 Name} \hspace{2cm} \textbf{Candidate 2 Name} \hspace{2cm} \textbf{Candidate 3 Name}\\ 
\vspace{4cm}
\textbf{Candidate 1 UID} \hspace{3cm} \textbf{Candidate 2 UID} \hspace{2cm} \textbf{Candidate 3 UID}\\ 
\vspace{1cm}

\pagebreak
\section*{Abstract}
Place your abstract here
\pagebreak
\section*{List of Figures}
Add your list of figures here
\pagebreak
\section*{List of Tables}
Add your list of tables here
\pagebreak
\tableofcontents
\pagebreak
\section{Introduction}
Note: Use subsections and subsubsections wherever required throughout the report
\newline Introduce your topic here
\subsection{Problem Statement}
\subsection{Objectives}
\subsection{Scope}
\subsection{Technologies Used}
\newline 
use subsections evrywhere Assumptions, constraints
\section{Literature Survey}
Note: Use subsections and subsubsections wherever required throughout the report
\newline Must contain a detailed study of minimum 12 research papers on the project topic and gap finding
\newline Use table captions to name tables at the top and use figure titles below figures wherever applicable 
\subsection{Literature Survey 1}
\subsection{Literature Survey 2}
...
\newline use page breaks after each section


\section{Analysis}
 Note: Use subsections and subsubsections wherever required throughout the report
\newline Use table captions to name tables at the top and use figure titles below figures wherever applicable 
Architecture/DFDs/class diagram/use-case diagram/sequence diagram/ flowchart/flow-diagram(others-as suggested by your guide)
\subsection{Analysis diagram 1}
\subsection{Analysis diagram 2}
\subsection{Analysis diagram 3}
.....
\newline Use page breaks after each section
\section{Design and Methodology}
 Note: Use subsections and subsubsections wherever required throughout the report
\newline Use table captions to name tables at the top and use figure titles below figures wherever applicable 
\newline Proposed System-draw the detailed block diagram to highlight the major modules in your system 
\newline Detail the algorithms/techniques(cite the algorithms/ models adopted from research papers here) you will be using for implementing each block in your proposed system
\newline Datasets used
\newline Add any other relevant details
\subsection{Design and Methodology 1}
\subsection{Design and Methodology 2}
\subsection{Design and Methodology 3}
...
\newline Use page breaks after each section
\section{Results and Discussion}
 Note: Use subsections and subsubsections wherever required throughout the report
\newline Write the details of implementation/results obtained/relevant screenshots/any relevant discussion and comparative analysis of methodologies used
Label all tables and figures wherever required
\subsection{Results and Discussion 1}
\subsection{Results and Discussion 2}
...
\newline Add any other relevant details
\newline Use page breaks after each section

\section{Conclusion and Further Work}
 Note: Use subsections and subsubsections wherever required throughout the report
summarise your conclusions and further work to be done for the next semester
\section{Research Paper/ Poster}
 Note: Use subsections and subsubsections wherever required throughout the report
include the paper with plagiarism report here, if accepted for publication, attach the acceptance receipt/certificate if any
\section{Plagiarism Report}
Put your plagiarism report here
\newline Use page breaks after each section

\pagebreak  
\begin{thebibliography}{1}
%---------------------------------------------------------

%------------------------------------------------------------------------
\bibitem{emu1}
S. Ben Saoud, B. Dagues, H. Schneider, M. Metz and J.C. Magiot, `Real Time Emulator of Static Converters/Electrical Machines Application to the Test of Control Unit,' \emph{Proceedings of the IEEE Symposium on Industrial Electronics},\hskip 1em plus 0.5em minus 0.4em\relax vol. 2, pp. 856-861, June 1996.
\bibitem{joep}
Joep Jacobs, Dirk Detjen, Claus-Ulrich Karipidis, and Rik W. De Doncker, `Rapid prototyping tools for power electronic systems: Demonstration with shunt active power filters,' \emph{IEEE Transactions on Power Electronics},\hskip 1em plus 0.5em minus 0.4em\relax vol. 19, no. 2, pp. 500-507, March 2004.
\bibitem{anto1}
Antonello Monti, Enrico Santi, Roger A. Dougal, and Marco Riva, `Rapid prototyping of digital controls for power electronics,' \emph{IEEE Transactions on Power Electronics},\hskip 1em plus 0.5em minus 0.4em\relax vol. 18, no. 3, pp. 915-923, May 2003.


%------------------------------------------------------------------------
\bibitem{blasko}
V. Kaura, and V. Blasko `Operation of a phase locked loop system under distorted utility conditions,' \emph{IEEE Transactions on Industrial Applications},\hskip 1em plus 0.5em minus 0.4em\relax vol. 33, no. 1, pp. 58-63, Feb. 1997.
\bibitem{mari}
Marian P. Kazmierkowski and Luigi Malesani, 'Current control techniques for three-phase voltage-source PWM converters: A survey,'
\emph{IEEE Transactions on Industry Applications},\hskip 1em plus 0.5em minus 0.4em\relax vol. 45, no. 5, pp. 691-702, Oct. 1998.

%------------------------------------------------------------------------
\bibitem{shin}
Shin-ichi Hamasaki Hamasaki and Atsuo Kawamura, `Improvement of current regulation of line-current-detection-type active filter based on deadbeat control,' \emph{IEEE Transactions on Industry Applications},\hskip 1em plus 0.5em minus 0.4em\relax vol. 39, no. 2, pp. 536-540, April 2003.
\bibitem{lcl3}
Marco Liserre, Frede Blaabjerg, and Steffan Hansen, `Design and Control of an LCL-Filter-Based Three-Phase Active Rectifier,' \emph{IEEE Transactions on Industry Applications},\hskip 1em plus 0.5em minus 0.4em\relax vol. 41, no. 5, pp. 1281-1291, Sept. 2005.

\end{thebibliography}
\end{document} 
 
